---
output: github_document
---

```{r, include = FALSE}
# knitr::opts_chunk$set(
#   collapse = TRUE,
#   comment = "#>",
#   fig.path = "man/figures/README-",
#   out.width = "100%"
# )
# Load the package to access functions if needed for examples (optional)
# library(forecastApp)
```

# forecastApp: A Modular Shiny Forecasting Application

`{forecastApp}` is an interactive Shiny application for time series forecasting. It provides a user-friendly interface to upload data, configure various forecasting models, run predictions, and evaluate results. Built with the `golem` framework, it emphasizes modularity and scalability.

## Features

- **Data Handling**: Upload your own daily time series data in CSV or Excel format.
- **Preprocessing**: Aggregate data (daily to weekly), handle missing values, apply transformations, and automatically split into training and testing sets.
- **Interactive Visualization**: Decompose time series, view autocorrelation plots (ACF/PACF), and analyze results with interactive graphs powered by `plotly`.
- **Modeling**: Configure and run a comprehensive suite of forecasting models:
  - ARIMA (AutoRegressive Integrated Moving Average)
  - ETS (Exponential Smoothing State Space Model)
  - TBATS (Trigonometric, Box-Cox, ARIMA, Trend, and Seasonal components)
  - Prophet (from Facebook)
  - XGBoost (eXtreme Gradient Boosting)
  - GAM (Generalized Additive Models)
  - Random Forest
  - NNETAR (Neural Network Time Series)
  The application supports hyperparameter tuning for some models (e.g., XGBoost, Random Forest) using `tidymodels` and time series cross-validation.
- **Evaluation**: Assess model performance with common accuracy metrics (e.g., MAE, RMSE, MAPE).
- **Session Management**: Save and load your application state, including data, configurations, and results, for continuity.
- **Reporting**: Generate dynamic HTML or PDF reports summarizing your forecast analysis.
- **Internationalization**: Supports multiple languages for a broader user base.

## Architecture

This application is structured as a robust and scalable R package using the **[Golem framework](https://engineering-shiny.org/golem.html)**. Its design promotes modularity, reusability, and maintainability through the extensive use of **Shiny Modules**.

Key architectural components include:

-   **Main Application (`app.R`, `R/app_ui.R`, `R/app_server.R`):**
    -   `app.R`: The entry point for launching the Shiny application.
    -   `R/app_ui.R`: Defines the overall user interface layout using `bslib::page_navbar`, organizing the application into distinct navigation panels (Data, Model, Forecast Results, Validation, About). It integrates the UI components of various Shiny modules.
    -   `R/app_server.R`: Manages the application's reactive state using a central `reactiveValues` object (`r`). It orchestrates the data flow between modules, calls the server-side logic of each module, handles model execution, metric calculation, session management, and report generation.
-   **Shiny Modules (`R/mod_*.R`):** Each significant feature or section of the application is encapsulated within its own Shiny module. This includes:
    -   `mod_data_input`: For uploading and selecting time series data.
    -   `mod_preprocess_controls`: For data aggregation, imputation, transformation, and train/test splitting.
    -   `mod_model_config`: For selecting and configuring parameters for each forecasting model.
    -   `mod_decomposition_plot`: For visualizing time series decomposition.
    -   `mod_results_plot`: For displaying the main forecast plots.
    -   `mod_results_table`: For presenting model performance metrics.
    -   `mod_extra_plots`: For additional diagnostic plots.
    -   `mod_model_summary`: For summarizing the configurations and outcomes of run models.
    -   `mod_validation`: For time series cross-validation.
    Each `mod_*.R` file typically contains both `*_ui` (User Interface) and `*_server` (Server Logic) functions for its specific functionality.
-   **Time Series Ecosystem:** The application leverages a rich ecosystem of R packages for time series analysis and machine learning, including `forecast`, `prophet`, `xgboost`, `mgcv`, `ranger`, `timetk`, `recipes`, `parsnip`, `workflows`, `tune`, `dials`, and `yardstick`.
-   **Internationalization (`inst/i18n/`):** Translation files (JSON) are used to support multiple languages, making the UI adaptable to different locales.
-   **Data Storage (`inst/extdata/`):** Default datasets and holiday files are stored here for easy access and demonstration.

This modular architecture ensures that the application is maintainable, testable, and extensible, allowing for easy addition of new models or features.

## Getting Started

### Installation

You can install the development version of `{forecastApp}` from GitHub. You will need the `remotes` package.

```{r}
# install.packages("remotes") # If you don't have remotes installed
# Replace <YOUR_GITHUB_USERNAME> with the actual username/organization
# remotes::install_github("<YOUR_GITHUB_USERNAME>/forecastApp")
```

**Dependencies:** The app requires several R packages available on CRAN. Some underlying packages (like `prophet` or `xgboost`) may have additional system dependencies (e.g., a C++ compiler, Python for Prophet). For installation from source, you may need Rtools (Windows) or Xcode Command Line Tools (macOS). Ensure you have these system dependencies installed if you encounter issues.

### Usage

Once the package is installed, you can run the application with a single command:

```{r, eval=FALSE}
library(forecastApp)
forecastApp::run_app()
```

### Development

This application is structured as a Golem package. To run the app in development mode, you can open the RStudio project and run the following script:

```{r, eval=FALSE}
source("dev/run_dev.R")
```

## Deployment

This application is deployed and publicly accessible on `shinyapps.io`:

[https://ismaguilera.shinyapps.io/forecastApp/](https://ismaguilera.shinyapps.io/forecastApp/)

`shinyapps.io` is a cloud-based platform provided by Posit (formerly RStudio) that allows R users to host, share, and scale their Shiny applications without needing to manage servers or infrastructure. It simplifies the process of making Shiny apps available to a wider audience.
